Study of the effects of magnetic braking on the lithium abundances of the Sun and solar-type stars II

### 2023-04-01 project/overleaf/paper2/mblithium.tex
Add paper 2

% Title of the paper, and the short title which is used in the headers.
% Keep the title short and informative.
\title[Rotation, magnetic braking \& Li abundances]{Study of the effects of magnetic braking on the lithium abundances of the Sun and solar-type stars}

% The list of authors, and the short list which is used in the headers.
% If you need two or more lines of authors, add an extra line using \newauthor
\author[R. Caballero Navarro et al.]{
R. Caballero Navarro,$^{1}$\thanks{E-mail: rcaballeron@correo.ugr.es}
A. Garc\'ia Hern\'andez,$^{1,2}$\thanks{E-mail: agh@ugr.es}
A. Ayala$^{2},$\thanks{E-mail: aayala@ugr.es}
J.~C. Su\'arez$^{1,2}$\thanks{E-mail: jcsuarez@ugr.es}
\\
% List of institutions
% Affiliations should be in the format ‘Department, Institution, Street
% Address, City and Postal Code, Country’.
$^{1}$Dept. Theoretical Physics and Cosmology, University of Granada (UGR), 18071, Granada, Spain\\
$^{2}$Instituto de Astrof\'isica de Andaluc\'ia (CSIC), Glorieta de la Astronom\'ia S/N, 18008, Granada, Spain\\
}

% These dates will be filled out by the publisher
\date{Accepted XXX. Received YYY; in original form ZZZ}

% Enter the current year, for the copyright statements etc.
\pubyear{2019}

@@

% Abstract of the paper
\begin{abstract}
\textit{Goals.} The study of lithium (Li) surface abundance in the Sun and young stellar globular clusters which are seemingly anomalous in present-day scenarios, as well as the influence of rotation and magnetic braking (MB) on its depletion during pre-main sequence (PMS) and main sequence (MS).
\newline\textit{Methods.} The effects of rotational mixing and of the rotational hydrostatic effects on Li abundances are studied by simulating several grids of PMS and MS rotating and non-rotating models. The previous effects are combined with the additional impact of the MB, (with magnetic field intensities ranging between 3.0 and 5.0 G). The data obtained from the two sets of simulations are confronted by comparing a number of parameters: Li abundances, angular velocity, effective temperature and size of convective zone.
\newline\textit{Results.} The surface Li abundance for the Sun like models at the end of the PMS and throughout the MS is decreased when rotational effects are included, i.e. the Li depletion rate for rotating models is higher than for non-rotating ones. However, this effect is attenuated when the braking effect produced by a magnetic field is present. This physical phenomena, generally neglected in stellar evolutionary models, impacts also the effective temperature ($\teff$) of the star and therefore its location in the HR diagram. The impact of MB in Li depletion is sensitive to the angular speed and this in turn on the magnetic field intensity: the higher the latter, the lower the Li destruction. A direct link between the magnetic fields and the convective zone (CZ) size is observed: stronger magnetic fields produce shallower CZ's. This result suggests that MB effect must be taken into consideration during PMS if we aim to reproduce Li abundances in young clusters.
\end{abstract}

@@

\section{Introduction} \label{sec_intro}
Despite decades of theoretical efforts, a coherent explanation for the Li abundance discrepancies detected in star clusters cannot be found. In young and old clusters these differences have been documented for stars in both PMS and MS  evolutionary stages. Additionally, these theoretical models are not able to explain the abundances detected in the late stages of the MS \citep{Tschape2001}.\par

Standard models, models that include convection only as a mixing process and do not consider other transport options like diffusion or angular momentum loss (AML), have been mainly involved in the elaboration of these predictions \citep{Sestito2005}. Li is destroyed at a temperature $\tli \approx 2.5 x 10^6\; K$ when a Li atom collides with a proton producing two He atoms. That takes place during proton-proton type II reactions (P-P II), and is therefore directly destroyed in stellar envelopes when the temperature at the base of the convection zone (BCZ) reaches  $\tli$. The Sun in particular and the solar-type stars in general are characterized by having a CZ that covers much of the stellar radius during the PMS which causes that their lower limit to exceed $\tli$ \citep{Iben1965}. This depletion stops during the approach to the zero-age MS (ZAMS) when the convection zone retreats and the temperature at BCZ is cooler than $\tli$. In the standard models only the mass and the initial chemical composition determine at what distance from the stellar surface $\tli$ temperature is reached. Therefore stars with similar mass belonging clusters are expected to reach the ZAMS with equal surface Li abundances. Furthermore, they should also show a very similar Li evolution until their approach to the terminal-age MS (TAMS). During this period the convection mechanisms unleash a mixing process that homogenizes the chemical composition of the envelope, from its lower limit till the star surface. However, different abundances of Li have been observed for different stellar populations \citep[see][and references therein]{Somers2014}.\par

Theoretical models take initial Li abundance as an input parameter and only describe how it is exhausted. Therefore, in order to make an accurate estimate of the initial abundance of Li, it is a prerequisite to be able to compare observations and models beforehand. Our Sun represents a unique exception, since it allows us to know the current abundance of this element in its photosphere, $A(\isotope[7]{Li}) = 1.1 \pm 0.1 \, dex$ \citep{Jeffries2004}, where $A(\isotope[7]{Li})$ is defined

\begin{ceqn}
\begin{align}
    A(\isotope[7]{Li}) &= log(N_{\isotope[7]{Li}} / N_{\isotope[1]{H}}) + 12
    \label{eq:A_Li}
\end{align}
\end{ceqn}

and its initial abundance of $A(\isotope[7]{Li}) = 3.34 \, dex$ is obtained from meteorite measurements \citep{Randich2006}. For other stars the initial Li abundance is found to be in the interval $3.0 \, dex < A(\isotope[7]{Li}) < 3.4 \, dex$. This has been estimated from  \citet{Randich2006}.\par

The impact of rotation both on PMS and Li depletion for solar-type stars has already extensively debated in the past \citep{Pinsonneault1997,Jeffries2004,Somers2014} and revised more recently on the basis of the availability of more accurate measures \citep{Bouvier2016}. However, these previous studies focused mainly on the hydrostatic effects and did not consider the influence of a coupling between the stellar magnetic field and its possible spin-down effect. AML has a direct influence in the mixing processes. It can be produced by a number of relevant mechanisms: mass loss, magnetic fields and gravity waves (g-modes). Gravity waves \citep{Charbonnel2005} and magnetic fields \citep{Eggenberger2009} transmit angular momentum (AM) much more effectively than inducing mixing \citep{Denissenkov2007}. As a consequence, the amount of differential rotation between the radiative and the convective zones of the star is reduced. Additionally, magnetic fields would also originate in the limits between these zones, in the so-called tachoclines \citep{Aschwanden2014, Guerrero2016}, which interact with the solar wind particles that are forced to co-rotate with it. This slows down the star, effect which is known as magnetic braking (MB). \par

Today, standard models are not able to replicate the observed values of Li abundance on stellar surfaces. This might happen because certain physical mechanisms that influence the destruction of Li are either being modelled improperly or simply not being taken into account, e.g. the magnetic braking. In this paper we study the interactions between rotation and magnetic fields when it comes to discuss the AM distribution. \par

\section{Method} \label{sec_method}

In the remainder of this paper, we describe a semi-empirical approach for the calculation of the AML as a result of the torque applied by a magnetically-coupled stellar winds. This has been implemented as an extension to Modules for Experiments in Stellar Astrophysics stellar evolution code \citep[MESA; ][]{Paxton2011, Paxton2013,Paxton2015, Paxton2018, Paxton2019}.

\subsection{Modelling magnetic braking} \label{mod_mb}
We start by enumerating the most relevant aspects and assumptions made in modelling the evolution of rotation, magnetic braking and angular momentum in MESA. These will be considered for the calculation of the AML as a result of the torque applied by a magnetically-coupled stellar winds.\par

A relevant parameter to characterize the influence of a given magnetic field on the stellar wind is the denominated wind confinement magnetic parameter ($\eta_*$). It represents a energy ratio (Eq.~\ref{eq:wind_conf}) and defines a characteristic parameter for the relative effectiveness of the magnetic fields in circumscribing and/or channeling the wind outflow \citep{UdDoula2002}.\par

\begin{ceqn}
\begin{equation}
    \eta_* = \frac{B^{2}/8\upi}{\rho\nu^2/2} \label{eq:wind_conf}
\end{equation}
\end{ceqn}

where $B^{2}/8\upi$ is the magnetic field energy density, $\frac{1}{2}\rho\nu^{2}$ the kinetic energy density, $B$ the magnetic field intensity, $\rho$ the mass density, and $\nu$ the stellar wind velocity.\par

Stars with similar initial masses but different mass loss ($\Dot{M}$) ratios will end up evolving very differently. The ionized particles carried by the solar wind not only contribute to the mass loss but also to the loss of kinetic ($K_e$) energy that is deposited in the interstellar medium. Given a star with a spherically symmetric wind, $\Dot{M}$ is characterized by the following expression:

\begin{ceqn}
\begin{equation}
    \Dot{M} = 4\upi r^2\rho\nu \label{eq:mass_loss}
\end{equation}
\end{ceqn}

Using (\ref{eq:wind_conf}) and (\ref{eq:mass_loss}), $\eta_*$ can be approximated by: 
\begin{ceqn}
\begin{equation}
    \eta_* = \frac{B^{2}r^{2}}{\Dot{M}\nu} \label{eq:wind_conf2}
\end{equation}
\end{ceqn}

In general, a magnetically channeled outflow has a complex stream geometry but for convenience, the Eq.~\ref{eq:mass_loss} simply characterizes the wind strength in terms of a spherically symmetric mass-loss rate. $\nu$ can be characterized by the radial variation of outflow velocity in terms of the velocity law (Eq.~\ref{eq:vel_law})
\begin{ceqn}
\begin{equation}
    \nu(r) = \nu_\infty (1-\rstar/r) \label{eq:vel_law}
\end{equation}
\end{ceqn}
where $r$ represents the distance from the center of the star to the point where the velocity is to be measured, $R_*$ is the radius of the star, and $\nu_\infty$ is the terminal wind velocity. $\nu_\infty$ is defined as the velocity that the wind reaches at large distance from the central star, where it is not accelerated anymore by the wind driving force but its deceleration due to interaction with the interstellar medium (ISM) is negligible \citep{Niedzielski2002}.\par

The line-driven winds of OB stars have terminal velocities that scale with the photospheric escape velocity \citep{Lamers2000} according to:
\begin{ceqn}
\begin{align}
    \nu_\infty &\simeq 1.92 \;\nu_{{\rm esc}} \label{eq:vinf}\\
    \nu_{{\rm esc}} &= \sqrt{\frac{2\,G\,\mstar}{\rstar}} \label{eq:vesc}
\end{align}
\end{ceqn}
where Eq.~\ref{eq:vesc} describes the Newtonian escape velocity from the stellar surface, where $G$ is the gravitational constant,   ${\rstar}$ the radius, and ${\mstar}$ the mass of the star.

Making use of (\ref{eq:vesc}), $\nu_\infty$ can be reformulated as:

\begin{ceqn}
\begin{equation}
    \nu_\infty \simeq 1.92 \; x \; 618 \; \Bigg(\sqrt{\frac{\rsun}{\rstar}\frac{\mstar}{\msun}} \;\Bigg) \label{eq:vinf}
\end{equation}
\end{ceqn}
where ${\rsun}$ is the radius, and ${\msun}$ the mass of the Sun.

It has been observed that strongly magnetic intermediate-mass stars typically have rotation rates much slower than other stars in their parent population \citep{Mathys2006}. In those stars, the magnetic fields interacts with the mass loss, where the Alfv\'{e}n radius ($R_{A}$) plays an important role. Making use of $\eta_*$ (Eq. \ref{eq:wind_conf}), $R_{A}$ is defined as the point in which the magnetic field energy density and the kinetic energy density are balanced. In case that $R_{A}$ is greater than the stellar radius, then the wind flow will have to follow the magnetic field. As a consequence, the material leaves the stellar surface with a higher specific AM, as the co-rotation radius has increased and it roughly corresponds to $R_{A}$. Following \citet{Weber1967} the AML can be calculated by:
\begin{ceqn}
\begin{equation}
 \Dot{J} = \frac{2}{3} \Dot{M}\Omega R^{2}_{A} \label{eq:j_dot}
\end{equation}
\end{ceqn}
where $\Dot{M}$ is the mass loss rate, $\Omega$ the angular velocity at star surface and $R_A$ the Alfv\'{e}n radius. \par

The above expression can be rewritten according to \citet{Ud-Doula2008} so that it depends on the $\eta_*$ instead of the $R_A$ as follows:
\begin{ceqn}
\begin{equation}
 \Dot{J} = \frac{2}{3} \Dot{M}\Omega R^{2}_{*}\eta_* \label{eq:j_dot_mesa}
\end{equation}
\end{ceqn}
which is a more convenient expression to be implemented in MESA because is based on values directly exposed during the simulations.

>>>>

% Title of the paper, and the short title which is used in the headers.
% Keep the title short and informative.
\title[Rotation, magnetic braking \& Li abundances II]{Study of the effects of magnetic braking on the lithium abundances of the Sun and solar-type stars II}

% The list of authors, and the short list which is used in the headers.
% If you need two or more lines of authors, add an extra line using \newauthor
\author[R. Caballero Navarro et al.]{
R. Caballero Navarro,$^{1}$\thanks{E-mail: rcaballeron@correo.ugr.es}
A. Garc\'ia Hern\'andez,$^{1,2}$\thanks{E-mail: agh@ugr.es}
J.~C. Su\'arez$^{1,2}$\thanks{E-mail: jcsuarez@ugr.es}
\\
% List of institutions
% Affiliations should be in the format ‘Department, Institution, Street
% Address, City and Postal Code, Country’.
$^{1}$Dept. Theoretical Physics and Cosmology, University of Granada (UGR), 18071, Granada, Spain\\
$^{2}$Instituto de Astrof\'isica de Andaluc\'ia (CSIC), Glorieta de la Astronom\'ia S/N, 18008, Granada, Spain\\
}

% These dates will be filled out by the publisher
\date{Accepted 2023 XXX XX. Received 2023 XXXX XX; in original form 2023 XXXX XX}

% Enter the current year, for the copyright statements etc.
\pubyear{2023}

@@

% Abstract of the paper
\begin{abstract}
The study of lithium (Li) surface abundance in the Sun and young stellar globular clusters which are seemingly anomalous in present-day scenarios, as well as the influence of rotation and magnetic braking (MB) on its depletion during pre-main sequence (PMS) and main sequence (MS).
In this work, the effects of rotational mixing and of the rotational hydrostatic effects on Li abundances are studied by simulating several grids of PMS and MS rotating models. Those effects are combined with the additional impact caused by the MB effect. Its intensity evolves as the star does i.e. not remaining fixed through the simulations. The data obtained from simulations are confronted with observational data measured on young open clusters collected by the Gaia-ESO Survey (GES). The results show how faster rotating models destroy more Li in the late PMS and during the MS, and how the effects of magnetic braking mitigate this destruction. The interconnection between the star's effective temperature, surface gravity and rotational velocity, as well as its chemical composition, determine the strength of the magnetic field and impact the effectiveness of the convective motions involved in the destruction of Li. The present work documents an approach in which neither the magnetic field strength nor the MLT parametrisation remains fixed throughout the evolution of the star, and which ends up providing results that are compatible with observational data for Li abundances of solar siblings hosted in young clusters.

\end{abstract}

@@

\section{Introduction} \label{sec_intro}
The angular momentum loss (AML) is considered to be governed by magnetically-coupled stellar winds which exert a braking force on the star that eventually causes a loss of angular momentum \citep{UdDoula2002}.

Over the last few years, decades, data from different observational campaigns have allowed studies of the rotational evolution of low and medium-mass stars \citep{Hartman2010,Gallet2013,Bouvier2016}. 
Among them we find studies \citep{Ud-Doula2008,Cranmer2011,Gallet2015,Amard2016} that focus on modelling the effects of magnetic braking (MB) and how its coupling with stellar winds impact the rotational evolution of the star. These models rely on a number of parameters which are fixed, among them the magnetic field strength B (we will come back to this point later). This approach is not sufficient to explain the observational data, and interaction with other processes in the angular velocity evolution of low-mass stars is necessary: the distribution of angular momentum loss in the interior of the star along the pre-main sequence (PMS) and main sequence (MS) \citep{Charbonnel2005, Eggenberger2008, Eggenberger2009, Navarro2020}, and the star - protostellar disc interaction during the PMS \citep{Bouvier2008,Gallet2013,Eggenberger2012,Zanni2012,Bouvier2016}.\par

As noted in our earlier work, discrepancies in Li abundances have been detected in young cluster for which there is as yet no coherent explanation \citep[see][and references therein]{Navarro2020}. Li is destroyed in stellar envelopes when the temperature at the base of the convection zone (BCZ) reaches $\tli \approx 2.5 x 10^6\; K$. The solar-type stars are known to have a convection zone (CZ) that covers much of the stellar radius during the PMS. In this way, the convective movements that occur in its interior manage to drag the Li to areas in which the BCZ limit exceeds the $\tli$ and Li is burned \citep{Iben1965}. This process slows down as the star approaches the zero-age MS (ZAMS) when the convection zone retreats and moves away from areas where the internal temperature is high enough to destroy the Li.\par

The impact of rotation both on PMS and Li depletion for solar-type stars has already extensively debated in the past \citep{Pinsonneault1997,Jeffries2004,Somers2014} and revised more recently on the basis of the availability of more accurate measures \citep{Gallet2013,Bouvier2016, Bouvier2018, Franciosini2022}. Of particular interest for this paper are the Li observational data measured on young open clusters. The Gaia-ESO Survey (GES) \citep{Gilmore2012,Randich2013,Randich2022} has produced very high quality spectroscopy records of around 110 000 Milky Way stars, in the field and in open clusters, down to magnitude 19, It has delivered good-quality radial velocities and stellar parameters for a large fraction of its unique target stars. Elemental abundances were derived for up to 31 elements for targets observed with the Fibre Large Array Multi Element Spectrograph (FLAMES) and the UV-Visual Echelle Spectrograph (UVES). Lithium abundances are delivered for about 1/3 of the sample.\par

As indicated by its name, GES intends to complement the data about parallaxes and proper motions from the Gaia satellite \citep{Mignard2005} with extremely precise information on radial velocities (RVs), Li, and chemistry in general. The first Gaia data release (GDR1) \citep{Brown2016} dated on 2016 and contained information on the first 14 months of operation. Gaia data release 2 (GDR2) \citep{Brown2018}, released on 2018, contained positions, parallaxes, and proper motions for about 1.3 billion sources, together with photometry. The third data release of Gaia has been divided in two parts: the Gaia early data release 3 (GEDR3) \citep{Brown2021} on 2020, and the Gaia data release 3 (GDR3) \citep{Brown2022} on 2022. They convey updated and more precise positions, parallaxes, proper motions, and photometry, spectra, classification and astrophysical parameters. The Gaia-ESO Survey will supplement the Gaia data for a significant subset of its targets, so that Gaia precision astrometry can be coupled with Gaia-ESO precision radial velocities and chemical abundances. In addition, a large fraction of the Gaia-ESO initiative is dedicated to the study of a large sample of open clusters (OCs).\par

GES collected data are very handy for our research because they conform a large data set of stellar parameters and Li abundances for several OCs, covering a large age range, from a few million to several billion years \citep{Dumont2021}. The analysis of surface Li abundances in young OCs of different ages can provide important constraints on stellar evolutionary models. As mentioned, GES enhances Gaia observation and several data releases has been made available. Of particular interests for this paper, it is the  latest release (DR5.0). The DR5.0 release includes all the astrophysical parameters derived by the Gaia-ESO consortium for 114 324 stars. In detail the catalogue provides \citep{Gilmore2022}:
\begin{itemize}
    \item The radial and projected RVs.
    \item Stellar parameters (effective temperature, surface gravity and metallicity).
    \item Abundances of several elements, among them Li.
    \item Specific parameters for tracing accretion and activity in young stars.
    \item Cluster probability membership.  
\end{itemize}

\section{Data} \label{sec_method}
The datasets provided by the Gaia mission in GEDR3 and GDR3, as well as those published by GES in DR5.0 form the main references used in this work. In general, we focus on the observations made on OCs, and particularly on those analysed in the \cite{Bragaglia2022} and \cite{Randich2022} papers. On the basis of the results presented in these papers, a set of 65 OCs are selected as the reference dataset for our work.

\begin{table}
	\centering
	\caption{Information on the pre-selected OCs.}
	\label{tab:oc_list}
	\begin{tabular}{l l l l} 
		\hline
		Cluster & GES\_FLD & age (Gyr) & [Fe/H]\\
		\hline
            25 Ori & 25\_Ori & 0.013 & 0 \\
            ASCC 50 & Assc50 & 0.011 & 0.02 \\
            Berkeley 21 & Br21 & 2.138 & -0.21 \\
            Berkeley 22 & Br22 & 2.455 & -0.26 \\
            Berkeley 25 & Br25 & 2.455 & -0.25 \\
            Berkeley 30 & Br30 & 0.295 & -0.13 \\
            Berkeley 31 & Br31 & 2.818 & -0.29 \\
            Berkeley 32 & Br32 & 4.898 & -0.31 \\
            Berkeley 36 & Br36 & 6.761 & -0.15 \\
            Berkeley 39 & Br39 & 5.623 & -0.14 \\
            Berkeley 44 & Br44 & 1.445 & 0.22 \\
            Berkeley 73 & Br73 & 1.413 & -0.26 \\
            Berkeley 75 & Br75 & 1.738 & -0.34 \\
            Berkeley 81 & Br81 & 1.148 & 0.22 \\
            Blanco 1 & Blanco1 & 0.105 & -0.03 \\
            Chamaeleon I & Cha\_I & 0.001 & -0.03 \\
            Collinder 197 & Col197 & 0.014 & 0.03 \\
            Collinder 69 & lam\_Ori & 0.013 & -0.09 \\
            Czernik 24 & Cz24 & 2.692 & -0.11 \\
            Czernik 30 & Cz30 & 2.884 & -0.31 \\
            ESO 92-05 & ESO92\_05 & 4.467 & -0.29 \\
            Haffner 10 & Haf10 & 3.802 & -0.1 \\
            IC 2391 & IC2391 & 0.029 & -0.06 \\
            IC 2602 & IC2602 & 0.036 & -0.06 \\
            IC 4665 & IC4665 & 0.033 & 0.01 \\
            Loden 165 & Loden165 & 3 &  \\
            Messier 67 & M67 & 3.981 & -0.02 \\
            NGC 2141 & NGC2141 & 1.862 & -0.04 \\
            NGC 2158 & NGC2158 & 1.549 & -0.15 \\
            NGC 2232 & NGC2232 & 0.018 & -0.03 \\
            NGC 2243 & NGC2243 & 4.365 & -0.45 \\
            NGC 2244 & NGC2244 & 0.004 & -0.04 \\
            NGC 2264 & NGC2264 & 0.003 & -0.1 \\
            NGC 2355 & NGC2355 & 1 & -0.13 \\
            NGC 2420 & NGC2420 & 1.698 & -0.15 \\
            NGC 2425 & NGC2425 & 2.399 & -0.13 \\
            NGC 2451A & NGC2451 & 0.035 & -0.08 \\
            NGC 2451B & NGC2451 & 0.041 & -0.02 \\
            NGC 2516 & NGC2516 & 0.24 & -0.04 \\
            NGC 2547 & NGC2457 & 0.032 & -0.03 \\
            NGC 3293 & NGC3293 & 0.01 & -0.08 \\
            NGC 3532 & NGC3532 & 0.398 & -0.01 \\
            NGC 3766 & NGC3766 & 0.023 & -0.12 \\
            NGC 4815 & NGC4815 & 0.372 & 0.08 \\
            NGC 6005 & NGC6005 & 1.259 & 0.22 \\
            NGC 6067 & NGC6067 & 0.126 & 0.03 \\
            NGC 6253 & NGC6253 & 3.246 & 0.33 \\
            NGC 6259 & NGC6259 & 0.269 & 0.18 \\
            NGC 6281 & NGC6281 & 0.513 & -0.04 \\
            NGC 6405 & NGC6405 & 0.035 & -0.02 \\
            NGC 6530 & NGC6530 & 0.002 & -0.02 \\
            NGC 6633 & NGC6633 & 0.692 & -0.03 \\
            NGC 6649 & NGC6649 & 0.071 & -0.08 \\
            NGC 6705 & NGC6705 & 0.309 & 0.03 \\
            NGC 6709 & NGC6709 & 0.191 & -0.02 \\
            NGC 6802 & NGC6802 & 0.661 & 0.14 \\
            Pismis 15 & Pismis15 & 0.871 & 0.02 \\
            Pismis 18 & Pismis18 & 0.575 & 0.14 \\
            Ruprecht 134 & Rup134 & 1.66 & 0.27 \\
            Trumpler 14 & Trumpler14 & 0.003 & -0.01 \\
            Trumpler 20 & Trumpler20 & 1.862 & 0.13 \\
            Trumpler 23 & Trumpler23 & 0.708 & 0.2 \\
            Trumpler 5 & Trumpler5 & 4.266 & -0.35 \\
            Gamma Vel & gamma2\_vel & 0.02 & -0.02 \\
            Rho Oph & Rho\_Oph & 0.001 & 0.03 \\
		\hline
	\end{tabular}
\end{table}

\subsection{The selection of clusters}
GES DR5.0 provides information on 114 324 stars, not all of which correspond to OCs, see \cite{Gilmore2022} for details. Of interest for the present work are those identified as belonging to OCs. For them we proceed to perform a first filtering on the GES\_TYPE field to identify those records that meet the condition:
\begin{itemize}
    \item GE\_CL: observed by GES, OC programme field
    \item GE\_SD\_OC: observed by GES, standard field OC
    \item AR\_CL: ESO Archive observation, OC programme field
    \item AR\_SD\_OC: Archive observation, standard field OC
\end{itemize}

As a result of this first filtering, the number of selected records is reduced to 43 299. On this subset of data we perform another filtering, this time aimed at selecting those that offer information on the following attributes:
\begin{itemize}
    \item Metallicity
    \item Li abundance
    \item GES field identifier
    \item Effective temperature
    \item Surface gravity
\end{itemize}

This second filtering reduces the population of stars to be considered to a total of 21 125 records. Finally, a third filtering is carried out to select those stars belonging to the OCs listed in Table \ref{tab:oc_list}, giving us a final study population of 20 913 stars distributed in 65 OCs.\par

\subsection{The selection of closest solar siblings}
Once we have pre-selected the stars belonging to the OCs listed in the previous chapter, our interest focus on identifying those that are most similar to our Sun i.e. solar siblings. Finding such solar twins is a difficult task since they are disseminated across our Galaxy. Finding them would help us to figure out the origin of our Sun and and gain information about what conditions prevailed in the OCs in which they were born.\par

Several attempts have been made to find solar siblings in the past \citep[see][and references therein]{Adibekyan2018}. In most of them, the search method is oriented to search the candidates based on its kinematic properties and then their metallicities and chemical abundances are compared with those of our Sun.\par

In our case we extend the selection criteria by adding additional constraints: the surface gravity and age of the star. These two new parameters represent a fine-tuning of the search process for sibling candidates to our Sun. We use the age of the OCs reported in Table \ref{tab:oc_list} and the data obtained from the set of simulations performed using Modules for Experiments in Stellar Astrophysics stellar evolution code \citep[MESA; ][]{Paxton2011, Paxton2013,Paxton2015, Paxton2018, Paxton2019}. Our models extend the investigations documented in \cite{Navarro2020} on the influence of rotation and magnetic braking on the star Li depletion during the pre-main sequence (PMS) and the main sequence (MS). The data obtained from the simulations are confronted by comparing different stellar parameters obtained by Gaia and GES. This allows us to analyse, validate and refine our models against a very large set of candidates. The end result of this process is to obtain a subset of candidate sibling stars of our Sun that help us to support (or refute) our line of investigation of how the Li abundance behaves throughout the evolution of solar-type stars.\par

From the approximately 21 000 identified stars, we proceed to select those with metallicity ($\feh$), effective temperature ($\teff$), and surface gravity ($\gsurf$) within the following limits:
\begin{itemize}
    \item $0.05 \, \dex <= \feh <= 0.05 \, \dex$
    \item $50.0 \, \Kelvin <= \teff <= 50.0 \, \Kelvin$
    \item $0.05 \, \dex <= \gsurf <= 0.05 \, \dex$
\end{itemize}

\section{Method} \label{sec_method}
The AML is considered to be governed by magnetically-coupled stellar winds which exert a braking force on the star that eventually causes a loss of angular momentum. In addition, throughout the star's evolution, directly observable parameters such as its rotational velocity, effective temperature, surface gravity or abundance of chemical elements undergo variations. These parameters have an influence, either directly or indirectly, on the angular momentum loss caused by the star's magnetic field. Therefore, assuming that the magnetic field strength of a star remains constant throughout its evolution is a necessary simplification in some scenarios but does not correspond to observations.\par

In the remainder of this paper, we describe a semi-empirical approach for the calculation of the AML as a result of the braking caused by a magnetic field of varying strength. The magnetic field evolves according to the $\teff$ and angular velocity ($\Omega$) of the star. This has been implemented as an extension to MESA.

\subsection{Modelling magnetic intensity} \label{mod_mbi}
The variable magnetic field strength modelling approach is supported by the papers \cite{Cranmer2011} and \cite{Gallet2013}. We start by enumerating the most relevant aspects and assumptions made in modelling the evolution of magnetic field intensity.\par

A relevant parameter to characterize the influence of a given magnetic field on the stellar wind is $\Omega$. Assuming that the magnetic field is generated by a dynamo, its strength is proportional to some power of $\Omega$,
\begin{ceqn}
\begin{equation}
    f_*B_* \propto \Omega_*^b \label{eq:mf_strenght}
\end{equation}
\end{ceqn}
where $b$ is the dynamo exponent, $B_*$ is the strength of the magnetic filed, and $f_*$ is the filling factor, i.e. the fraction of the star surface which is magnetized \citep{Gallet2013}.

On the other hand, $B_*$ is proportional to the equipartition magnetic field strength,
\begin{ceqn}
\begin{equation}
    B_* \approx 1.13 \Beq \label{eq:mf_bstar}
\end{equation}
\end{ceqn}
where $\Beq$ is defined as 
\begin{ceqn}
\begin{equation}
    \Beq = \sqrt{8\upi P_*} = \sqrt{\frac{8\upi\rho_* \boltzmann \teff}{\mu\massH}}\label{eq:mf_beq}    
\end{equation}
\end{ceqn}
where $\rho_*$ is the photospheric density, $\boltzmann$ is Boltzmann's constant, $\massH$ is the mass of a hydrogen atom, and $\mu$ the mean atomic weight \citep{Cranmer2011}.

As indicated in \cite{Cranmer2011} $\mu$ can be estimated using the OPAL plasma equations of state\footnote{\url{https://opalopacity.llnl.gov/EOS_2005/}},
\begin{ceqn}
\begin{equation}
    \mu \approx \frac{7}{4} + \frac{1}{2} \; {\rm tanh}\Bigg(\frac{3500-\teff}{600}\;\Bigg) \label{eq:mf_atom_weight}
\end{equation}
\end{ceqn}
Once we are in a position to calculate $\Beq$, we can calculate $B_*$.\par

\cite{Cranmer2011} found that the magnetic field strength is mainly influenced by $f_*$ rather than by the rotation period of the star. That is, variations in the star's angular velocity do not significantly alter its magnetic field strength. In addition, $f_*$ has a strong dependence on the Rossby number ($\rossby$). In order to calculate $\rossby$ of a given star, it is necessary to know the turn over time ($\turnover$),
\begin{ceqn}
\begin{equation}
    \rossby = \frac{\prot}{\turnover} \label{eq:mf_rossby}
\end{equation}
\begin{ceqn}
\begin{equation}
    \turnover = 314.24\;exp\left[-\Bigg(\frac{\teff}{1952.5}\Bigg)-\Bigg(\frac{\teff}{6250}\Bigg)^{18} \;\right]+0.002 \label{eq:mf_turnover}
\end{equation}

\end{ceqn}
\cite{Cranmer2011} Crammer provides two different settings for $f_*$ which define respectively the upper and lower limits for the mean surface magnetic field strength ($f_*B_*$),

\begin{ceqn}
\begin{align}
     \fmin &= \frac{0.5}{(1+(x/0.16)^{2.6})^{1.3}} \label{eq:fmin}\\
     \fmax &= \frac{1}{1+(x/0.31)^{2.5}} \label{eq:fmax}
\end{align}
\end{ceqn}
where x = $\rossby$/$\rossbysun$, and $\rossbysun$ = 1.96.

Finally, we are able to calculate the filling factor for a star by applying the adjustment made by \cite{Gallet2013},
\begin{ceqn}
\begin{equation}
     \fstar = \frac{0.55}{(1+(x/0.16)^{2.3})^{1.22}} \label{eq:fstart}
\end{equation}
\end{ceqn}

\begin{figure}
	\includegraphics[clip,width=\columnwidth]{figures/paper2/mag_field_limits_var_vel_g3.pdf}
    \caption{The evolution of magnetic field intensity and its upper and lower limits, as a function of time and $\oomegac$ for several 1 $\msun$ models and their. The model include PMS rotation with $\oomegac$ = 0.1425. The solid lines represent the magnetic field strength, while the dotted lines represent the angular evolution of the star. The dashed vertical line makes reference to the ZAMS.}
    \label{fig:mag_field_limits_var_vel_g3}
\end{figure}

With this formalism we are able to calculate the mean magnetic field as a time-dependent function of $\rho$, $\teff$, and $\Omega$. The time values for these three variables will be obtained from the different simulations carried out with the help of MESA.

\subsection{Modelling magnetic braking} \label{mod_mb}
Once we have characterised how the magnetic field strength evolves, we need to model how the loss of angular momentum occurs.  In \cite{Navarro2020} we use a model applicable to magnetic fields of non-variable intensity and of few Gauss ($1.0\,\Gauss \leq B \leq 10.0\,\Gauss$). As shown in Figure \ref{fig:mag_field_limits_var_vel_g3}, the simulations yield magnetic field values that show maximum intensities of around $1.0K\,\Gauss$, making the use of the model employed previously non viable.

Let's start by enumerating the most relevant aspects and assumptions made in modelling the evolution of rotation, magnetic braking and angular momentum. If we assume a spherical outflow, the torque applied by a magnetically-coupled stellar winds $\torquewind$ is dictated by,
\begin{ceqn}
\begin{equation}
    \torquewind \propto \Omega_* \; \mwind \; \ralfven^{2} \label{eq:mb_torque}
\end{equation}
\end{ceqn}
where $\mwind$ is the mass loss rate, and $\ralfven$ the averaged value of the Alfvén radius.\par
Stars with similar initial masses but different mass loss ($\Dot{M}$) ratios will end up evolving very differently. The ionized particles carried by the solar wind not only contribute to the mass loss but also to the loss of kinetic energy that is deposited in the interstellar medium. Given a star with a spherically symmetric wind, $\mwind$ is characterized by the following expression:

\begin{ceqn}
\begin{equation}
    \mwind = 4\upi r^2\rho\nu \label{eq:mass_loss}
\end{equation}
\end{ceqn}
where $r$ is the stellar radius, $\rho$ the mass density, and $\nu$ the stellar wind velocity.
